### 2011-03-20 extra/sliver_pldi11/klimited.tex
first pass

% Basics
We now apply our general pruning approach to $k$-limited pointer analysis.
Each node in the control-flow graph of each method $m \in \M$ is associated with a
simple statement (e.g., $v_2 = v_1$).
We omit statements that have no effect on our analysis (e.g., operations on
data of primitive type).
For simplicity, we assume each method has a single argument and no return value.\footnote{
Our actual implementation is a straightforward extension of this simplified analysis
which handles multiple arguments, return values, class initializers, and
objects allocated through reflection.}
\reffig{aliasAnalysisRules} describes the Datalog program corresponding to this
analysis.

% Context, abstraction
Our analysis represents both contexts and abstract objects using a list of
allocation sites, which we will call {\em chains}.  However, note that these
chains are never truncated in the Datalog program.  Therefore, running the
Datalog program (ignoring the fact that it might not terminate) is performing
an $\infty$-object-sensitivity analysis.  Recall that we separate the
computation of the query in Datalog with the abstraction which is applied to
the input tuples (of course, our Datalog program is itself an approximation to
the concrete semantics---it is flow-insensitive, does not handle primitive
data, etc.).  \refsec{abstractions} will describe further abstractions of this
analysis.

@@

We briefly describe the analysis rules in Datalog.  Rule (1)
states that the main method $\mmain$ is reachable in a distinguished context $[]$.
Rule (2) states that a target method of a reachable call site is also reachable.
Rule (3) states that every statement in a reachable method is also reachable.
Rules (4) through (9) implement the transfer function associated with each kind of
statement.
%Rule (10) populate the call graph and
%rule (11) propagate the points-to set from the argument
%of a call site to the formal argument of each target method.
Rule (10) analyzes the target method
$m$ in a separate context $o$ for each abstract object $o$ to which the distinguished
{\tt this} argument of method $m$ points, and rule (11) sets the points-to set
of the {\tt this} argument of method $m$ in context $o$ to the singleton $\{ o \}$.

\begin{figure*}
\begin{minipage}{3.3in}
{\bf Domains}: \\
\[
\begin{array}{rrcl}
\mbox{{(method)}}                  & m   &\in & \M = \{ \mmain, ... \}\\
\mbox{{(local variable)}}          & v   &\in & \V \\
\mbox{{(global variable)}}         & g   &\in & \G  \\
\mbox{{(object field)}}            & f   &\in & \F  \\
\mbox{{(method call site)}}        & i   &\in & \I \\
\mbox{{(allocation site)}}         & h   &\in & \H \\
\mbox{{(statement)}}               & p   &\in & \P \\
\mbox{{(method context)}}          & c   &\in& \C = \H^* \\
\mbox{{(abstract object)}}         & o   &\in& \O = \H^* \\
\end{array}
\]
\[
\begin{array}{ll}
p ::= & v = \new\,h \bigmid v_2 = v_1 \bigmid g = v \bigmid v = g \bigmid \\
      & v_2.f = v_1 \bigmid v_2 = v_1.f  \bigmid i(v)
\end{array}
\]
\end{minipage}
\begin{minipage}{3.7in}
{\bf Input relations}: \\
\[ \begin{array}{lcll}
\body     &\subset& \M \times \P & \text{(method contains statement)} \\
\trgt     &\subset& \I \times \M & \text{(call site resolves to method)} \\
\argI     &\subset& \I \times \V & \text{(call site's argument variable)} \\
\argM     &\subset& \M \times \V & \text{(method's formal argument variable)} \\
\ext      &\subset& \S \times \C \times \C & \text{(extend context with site)} \\
          & =     & \multicolumn{2}{l}{\{ (h, c, [h]+c) : h\in\H, c\in\C \}} \\
\end{array}
\]
{\bf Output relations}: \\
\[ \begin{array}{lcll}
\reachM   &\subset& \C \times \M & \text{(reachable methods)} \\
\reachP   &\subset& \C \times \P & \text{(reachable statements)} \\
\ptsV     &\subset& \C \times \V \times \O & \text{(points-to sets of local variables)} \\
\ptsG     &\subset& \G \times \O & \text{(points-to sets of static fields)} \\
\heap     &\subset& \O \times \F \times \O & \text{(heap graph)} \\
\cg       &\subset& \C \times \I \times \C \times \M & \text{(call graph)}
\end{array}
\]
\end{minipage}
{\bf Rules}: \\
\[
\begin{array}{lclr}
\reachM([],\mmain). &           &  & (1) \\
\reachM(c,m)              &\dlogUpdate& \cg(*,*,c,m). & (2) \\
\reachP(c,p)              &\dlogUpdate& \reachM(c,m),\ \body(m,p). & (3)
\\[8pt]
\ptsV(c,v,o)              &\dlogUpdate& \reachP(c, v = \new\,h),\ \ext(h, c, o). & (4) \\
\ptsV(c,v_2,o)            &\dlogUpdate& \reachP(c, v_2 = v_1),\ \ptsV(c,v_1,o). & (5) \\
\ptsG(g,o)                &\dlogUpdate& \reachP(c, g = v),\ \ptsV(c,v,o). & (6) \\
\ptsV(c,v,o)              &\dlogUpdate& \reachP(c, v = g),\ \ptsG(g,o). & (7) \\
\heap(o_2,f,o_1)          &\dlogUpdate& \reachP(c, v_2.f = v_1),\ \ptsV(c,v_1,o_1),\ \ptsV(c,v_2,o_2). & (8) \\
\ptsV(c,v_2,o_2)          &\dlogUpdate& \reachP(c, v_2 = v_1.f),\ \ptsV(c,v_1,o_1),\ \heap(o_1,f,o_2). & (9)
\\[8pt]
\cg(c,i,o,m)            &\dlogUpdate& \reachP(c,i),\ \trgt(i,m),\ \argI(i,v),\ \ptsV(c,v,o). & (10) \\
\ptsV(c,v,c)             &\dlogUpdate& \reachM(c,m),\ \argM(m,v). & (11)
\end{array}
\]
%\centering
\caption{Datalog implementation of our $k$-object-sensitivity points-to analysis with call-graph construction.
Our abstraction $\ba$ affects the analysis solely through $\ext$,
which specifies that when we prepend $s$ to $c$, we truncate the resulting sequence to length $\ba_{s}$.
}
\label{fig:aliasAnalysisRules}
\end{figure*}

@@

The core pointer analysis is used by three clients, which each define a set of queries.

@@

>>>>

% Basics
We now introduce our $k$-limited pointer analysis,
on which we will apply the Prune-Refine algorithm.
Each node in the control-flow graph of each method $m \in \M$ is associated with a
simple statement (e.g., $v_2 = v_1$).
We omit statements that have no effect on our analysis (e.g., operations on
data of primitive type).
For simplicity, we assume each method has a single argument and no return value.\footnote{
Our actual implementation is a straightforward extension of this simplified analysis
which handles multiple arguments, return values, class initializers, and
objects allocated through reflection.}
\reffig{aliasAnalysisRules} describes the Datalog program corresponding to this
analysis.

% Context, abstraction
The analysis represents both contexts and abstract objects using chains of
allocation sites ($\C = \H^*$).  Contexts are extended into new contexts via the $\ext$ relation,
which prepends an allocation site to a chain (e.g., $\ext(3,[12],[312])$).
Note that these chains are not truncated in the Datalog program, and therefore,
running the Datalog program directly (ignoring the fact that it might not
terminate) corresponds to performing an $\infty$-object-sensitivity analysis.

Although this Datalog program itself is an approximation to the concrete
program semantics---it is flow-insensitive, does not handle primitive data,
etc., we will informally say that a client query computed with respect to this
analysis yields a concrete answer.  In contrast, we obtain an abstract
answer by computing the client query with respect to a $k$-limited abstraction,
which we will discuss in \refsec{generalRelationship}.

% Rules
We now briefly describe the Datalog rules in \reffig{aliasAnalysisRules}.  Rule (1)
states that the main method $\mmain$ is reachable in a distinguished context $[\,]$.
Rule (2) states that a target method of a reachable call site is also reachable.
Rule (3) states that every statement in a reachable method is also reachable.
Rules (4) through (9) implement the transfer function associated with each kind of statement.
Rule (10) analyzes the target method
$m$ in a separate context $o$ for each abstract object $o$ to which the distinguished
{\tt this} argument of method $m$ points, and rule (11) sets the points-to set
of the {\tt this} argument of method $m$ in context $o$ to the singleton $\{ o \}$.

@@

\begin{figure*}
\begin{minipage}{3.3in}
{\bf Domains}: \\
\[
\begin{array}{rrcl}
\mbox{{(method)}}                  & m   &\in & \M = \{ \mmain, ... \}\\
\mbox{{(local variable)}}          & v   &\in & \V \\
\mbox{{(global variable)}}         & g   &\in & \G  \\
\mbox{{(object field)}}            & f   &\in & \F  \\
\mbox{{(method call site)}}        & i   &\in & \I \\
\mbox{{(allocation site)}}         & h   &\in & \H \\
\mbox{{(statement)}}               & p   &\in & \P \\
\mbox{{(method context)}}          & c   &\in& \C = \H^* \\
\mbox{{(abstract object)}}         & o   &\in& \O = \H^* \\
\end{array}
\]
\[
\begin{array}{ll}
p ::= & v = \new\,h \bigmid v_2 = v_1 \bigmid g = v \bigmid v = g \bigmid \\
      & v_2.f = v_1 \bigmid v_2 = v_1.f  \bigmid i(v)
\end{array}
\]
\end{minipage}
\begin{minipage}{3.7in}
{\bf Input relations}: \\
\[ \begin{array}{lcll}
\body     &\subset& \M \times \P & \text{(method contains statement)} \\
\trgt     &\subset& \I \times \M & \text{(call site resolves to method)} \\
\argI     &\subset& \I \times \V & \text{(call site's argument variable)} \\
\argM     &\subset& \M \times \V & \text{(method's formal argument variable)} \\
\ext      &\subset& \H \times \C \times \C & \text{(extend context with site)} \\
          & =     & \multicolumn{2}{l}{\{ (h, c, [h]+c) : h\in\H, c\in\C \}} \\
\end{array}
\]
{\bf Output relations}: \\
\[ \begin{array}{lcll}
\reachM   &\subset& \C \times \M & \text{(reachable methods)} \\
\reachP   &\subset& \C \times \P & \text{(reachable statements)} \\
\ptsV     &\subset& \C \times \V \times \O & \text{(points-to sets of local variables)} \\
\ptsG     &\subset& \G \times \O & \text{(points-to sets of static fields)} \\
\heap     &\subset& \O \times \F \times \O & \text{(heap graph)} \\
\cg       &\subset& \C \times \I \times \C \times \M & \text{(call graph)}
\end{array}
\]
\end{minipage}
{\bf Rules}: \\
\[
\begin{array}{lclr}
\reachM([],\mmain). &           &  & (1) \\
\reachM(c,m)              &\dlogUpdate& \cg(*,*,c,m). & (2) \\
\reachP(c,p)              &\dlogUpdate& \reachM(c,m),\ \body(m,p). & (3)
\\[8pt]
\ptsV(c,v,o)              &\dlogUpdate& \reachP(c, v = \new\,h),\ \ext(h, c, o). & (4) \\
\ptsV(c,v_2,o)            &\dlogUpdate& \reachP(c, v_2 = v_1),\ \ptsV(c,v_1,o). & (5) \\
\ptsG(g,o)                &\dlogUpdate& \reachP(c, g = v),\ \ptsV(c,v,o). & (6) \\
\ptsV(c,v,o)              &\dlogUpdate& \reachP(c, v = g),\ \ptsG(g,o). & (7) \\
\heap(o_2,f,o_1)          &\dlogUpdate& \reachP(c, v_2.f = v_1),\ \ptsV(c,v_1,o_1),\ \ptsV(c,v_2,o_2). & (8) \\
\ptsV(c,v_2,o_2)          &\dlogUpdate& \reachP(c, v_2 = v_1.f),\ \ptsV(c,v_1,o_1),\ \heap(o_1,f,o_2). & (9)
\\[8pt]
\cg(c,i,o,m)            &\dlogUpdate& \reachP(c,i),\ \trgt(i,m),\ \argI(i,v),\ \ptsV(c,v,o). & (10) \\
\ptsV(c,v,c)             &\dlogUpdate& \reachM(c,m),\ \argM(m,v). & (11)
\end{array}
\]
%\centering
\caption{Datalog implementation of our $k$-object-sensitivity points-to analysis with call-graph construction.
Our abstraction $\ba$ affects the analysis solely through $\ext$,
which specifies that when we prepend $s$ to $c$, we truncate the resulting sequence to length $\ba_{s}$.
}
\label{fig:aliasAnalysisRules}
\end{figure*}

@@

The core pointer analysis just described is used by three clients, which each
define a set of queries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{generalRelationship}{Relationship to general notation}

% Relate 
We now describe our analysis in \reffig{aliasAnalysisRules} in terms of our general notation presented in \refsec{preliminaries}.
The set of concrete values $\sC$ is the union of all the domains
(e.g.., allocation sites $\H = \{ 1, 2, 3, \dots \}$, abstract objects $\C = \H^*$, etc.).
The input tuples $X$ are specified by the input relations
(e.g., $X = \{ \body(\mmain, x = \new\,3), \ext(3,[12],[312]) \}$).
Each of the three clients defines a set of possible query tuples, for example,
$\xo = \dl{unsafe}(\dl{v4},\dl{v8})$ for downcast safety checking of an assignment $\dl{v4} = \dl{v8}$.

Recall that $\bP(X)$ corresponds to obtaining an answer to a client query with respect to $\infty$-object-sensitivity.
To obtain $k$-object-sensitivity,
we first apply the $k$-limited abstraction $\pi_k$ to the input tuples
and run the Datalog program on these abstract tuples ($\bP(\pi_k(X))$).
Note that only the $\ext$ tuples are affected by the abstraction.

### 2011-03-20 extra/sliver_pldi11/related.tex
first pass

% Larger space of abstractions
There is a wealth of literature which attempts to scale static analyses without
sacrificing precision.  One general theme is to work with a flexible family of
abstractions, which in principle allows us to conform to the needs of the
client.  Milanova et al.~\cite{kobj,MilanovaRountevRyder2005} also consider
abstractions, where each local variable can be independently treated context
sensitively or insensitively, and different $k$ values can be chosen for
different allocation sites.  Lhot\'{a}k and
Hendren~\cite{LhotakHendren2006,LhotakHendren2008} present Paddle, a
parametrized framework for BDD-based, $k$-limited alias analysis.
\cite{smaragdakis11context} scale up $k$-object-sensitivity one more level by
using types rather than allocation sites.  However, which parts of the
abstraction should be more refined is largely left up to the user.

% Client-driven
Client-driven approaches use feedback from a client query to determine what
parts of an abstraction to refine.
Plevyak and Chien~\cite{PlevyakChien1994} use a refinement-based algorithm for
type inference, where context sensitivity is driven by detecting type conflict.
Guyer and Lin~\cite{GuyerLin2003} present a pointer analysis for C
which detects loss of precision (e.g., at merge points) and introduce context sensitivity.
Our method for determining relevant input tuples is similar in spirit but more general.

Perhaps the closest to this work is the part of Liang et al.~\cite{liang11minimal}, which
uses the same technique to compute the set of relevant tuples.  However, what
is done with this information is quite different.  \cite{liang11minimal} merely
stops refining the irrelevant sites whereas we actually prune all irrelevant
tuples, thereby exploiting this information fuller.  As we saw in
\refsec{experiments}, this difference had major ramifications.
%An importance difference is that we are not

@@

% Demand-driven
Demand-driven analyses \cite{HeintzeTardieu2001,ZhengRugina2008} do not refine
the abstraction but rather compute an analysis on an existing
abstraction faster.  Some approaches such as \cite{SridharanBodik2006} do both.
Our \PR\ algorithm has a flavor of the client-driven approach, in that we do
refine our abstraction, but also of the demand-driven approach, in that we do
not perform a full computation (in particular, ignoring tuples which were
pruned).

@@

% Dynamic - lower bound
Pruning has been implemented in various settings.
\cite{vipindeep05pruning} use dynamic analysis to prune down set of paths and
then focus a static analysis on these paths.
\cite{spoon04ddp} use pruning for type inference in functional languages,
where pruning is simply a heuristic which shortcuts a search algorithm.
As a result, pruning can hurt precision.
One advantage of our pruning approach is that it comes with strong soundness
and completeness guarantees.

>>>>

% Larger space of abstractions
There is a wealth of literature which attempts to scale static analyses without
sacrificing precision.  One general theme is to work with a flexible family of
abstractions, which in principle allows us to conform to the needs of the
client.  Milanova et al.~\cite{kobj,MilanovaRountevRyder2005} consider
abstractions where each local variable can be independently treated
context-sensitively or context-insensitively, and different $k$ values can be
chosen for different allocation sites.  Lhot\'{a}k and
Hendren~\cite{LhotakHendren2006,LhotakHendren2008} present Paddle, a
parametrized framework for BDD-based, $k$-limited alias analysis.
\cite{smaragdakis11context} scale up $k$-object-sensitivity, increasing $k$ by one
using types rather than allocation sites.  However, in all of this work, which parts of the
abstraction should be more refined is largely left up to the user.

% Client-driven
Client-driven approaches use feedback from a client query to determine what
parts of an abstraction to refine.
Plevyak and Chien~\cite{PlevyakChien1994} use a refinement-based algorithm for
type inference, where context-sensitivity is driven by detecting type conflicts.
Guyer and Lin~\cite{GuyerLin2003} present a pointer analysis for C
which detects loss of precision (e.g., at merge points) and introduce context-sensitivity.
Our method for determining relevant input tuples is similar in spirit but more general.

Section~3.1 of Liang et al.~\cite{liang11minimal} (not the focus of that work)
also computes the set of relevant tuples by running a transformed Datalog program.  However, what
is done with this information is quite different.  \cite{liang11minimal} merely
stops refining the irrelevant sites whereas we actually prune all irrelevant
tuples, thereby exploiting this information fully.  As we saw in
\refsec{experiments}, this difference had major ramifications.

@@

% Demand-driven
Demand-driven analyses \cite{HeintzeTardieu2001,ZhengRugina2008} do not refine
the abstraction but rather try to compute an analysis on an existing
abstraction more efficiently.
Sridharan et al.~\cite{SridharanBodik2006} presents an algorithm
which casts pointer analysis as a CFL-reachability problem and relaxes by
introducing additional edges.

Our Prune-Refine algorithm has a flavor of the client-driven approach, in that we do
refine our abstraction, but also of the demand-driven approach, in that we do
not perform a full computation (in particular, ignoring tuples which were
pruned).  There are three important differences between the present work
and the work described earlier:
First, while most of that work is specific to pointer analysis, our Prune-Refine algorithm
is generally applicable to any Datalog program.
Second, while past work is based on selected iterative refinement,
Prune-Refinement is based on generalizing the idea of program slicing.
%Finally, we were able to prove rigorous theoretical guarantees that our algorithm is as precise
%as a more expensive algorithm.

@@

% Dynamic - lower bound
Other forms of pruning have been implemented in various settings.
\cite{vipindeep05pruning} use dynamic analysis to prune down set of paths and
then focus a static analysis on these paths.
\cite{spoon04ddp} use pruning for type inference in functional languages,
where pruning is simply a heuristic which shortcuts a search algorithm.
As a result, pruning can hurt precision.
One advantage of our pruning approach is that it comes with strong soundness
and completeness guarantees.

### 2011-03-20 extra/sliver_pldi11/theory.tex
first pass

%In this section, we show how we can use pruning to obtain a sound analysis which
%is complete with respect to a given abstraction.
We first describe the core
idea behind pruning (\refsec{oneStep}) and then show how it can be used in the
Prune-Refine (\PR) algorithm (\refsec{algorithm}).

@@

The left-hand side of \refeqn{master} corresponds to running the analysis with respect to $\alpha$.
The right-hand side corresponds to first pruning the input tuples $X$
with $\beta$ and then running the analysis with $\alpha$.
The theorem states that the two procedures obtain identical results.
The significance of this is that the right-hand side is often a much cheaper
way to compute the left-hand side.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{algorithm}{The Prune-Refine (\PR) algorithm}

\begin{figure}[t]
\begin{center} \scalebox{0.95}{\framebox{ \begin{minipage}{3.3in} % Begin model
\begin{center} Prune-Refine (\PR) Algorithm \end{center}
{\bf Input}: \\
\ind Sequence of abstractions: $\alpha_0 \preceq \alpha_1 \preceq \alpha_2 \preceq \cdots$ \\
\ind [Auxiliary abstractions: $\beta_t \preceq \alpha_t, t = 0, 1, 2, \dots$] \\
\ind $A_0 = \alpha_0(X)$, set of tuples \\
\\
For $t = 0, 1, 2, \dots$: \\
\ind [{\bf Pre-prune}: $A_t' \leftarrow A_t \cap \alpha_t(\bP(\beta_t(A_t)))$] \\
\ind {\bf Prune}: $\tilde A_t = \bP(A_t')$. If $\tilde A_t = \emptyset$: return {\em proven}. \\
\ind {\bf Refine}: $A_{t+1} = \alpha_{t+1}(\tilde A_t)$. \\
%\ind\ind\ind\ind\ind If $A_{t+1} = A_t$: return {\em not proven}.
\end{minipage} }} \end{center} % End model
\includegraphics[scale=0.5]{figures/algorithm}
\caption{\label{fig:pseudocode} The pseudocode and the schema for the \PR\ algorithm.
The algorithm maintains a set of (abstract) input tuples which could be
involved in some derivation of the query $\xo$ and attempts to prune down this set.
The basic version, which excludes parts in square brackets,
simply alternates between pruning and refining.  The full version includes
a pre-pruning step using auxiliary abstractions in order to further reduce the
number of tuples.
}
\end{figure}

We now turn \refthm{master} into a full algorithm, which we call the
Prune-Refine (\PR) algorithm.  \reffig{pseudocode} shows the pseudocode
of the algorithm and a diagram showing the computation of the various abstract
input tuples computed.

% Basic
This algorithm essentially applies \refeqn{master} repeatedly.
We first present a simplified version of the algorithm which
ignores the pre-pruning step (we take $A_t = A_t'$).
We are given a sequence of successively finer abstractions $\alpha_0, \alpha_1, \dots$,
and an input set of abstract tuples $A_0$, which is computed
under the initial abstraction $\alpha_0$.  Then the algorithm alternates between a pruning step
and a refining step, always maintaining only the abstract tuples
that could participate in a derivation of the query tuple $\xo$.
On iteration $t$, our current input tuples $A_t$ are first pruned to $\tilde A_t$ using $\bP$;
this is subsequently refined to $A_{t+1}$.
% Example
\reffig{graphIters} shows an example of running this algorithm on the graph
example from \reffig{graphExample}; the first pruning step is shown in
\reffig{graphDerivation}.

\Fig{figures/graphIters}{0.3}{graphIters}{
The abstract input tuples computed by
the \PR\ algorithm on the graph example from \reffig{graphExample} (without pre-pruning),
where the abstraction $\alpha_t$ maps $c$ to all paths that match $c$ on the first $t+1$ nodes.
Notation: $c*$ denotes the set of all paths with prefix $c$.
During the first pruning step, $\ext(2,\vec{0}*,\vec{2}*)$ is pruned from $A_0$,
yielding $\tilde A_0$.  In the refine step, we expand $\vec{1}*$ to $\{ \vec{1} \}$ and $\vec{10}*$,
resulting in an extra tuple.  In the second pruning step, we prove the query
(pruning everything).
}

% Pre-pruning
Now we discuss pre-pruning.  Pre-pruning requires the user to provide
another sequence of successively finer abstractions $\beta_0, \beta_1, \dots$
which are coarser than $\alpha_0, \alpha_1, \dots$, respectively.
These abstractions will also be used to prune the input tuples.
The idea is that before refining $A_t$ to $A_{t+1}$,
we perform two steps of pruning:
(i) first using $\beta_t$ during pre-pruning step, and then
(ii) using $\alpha_t$ during the main pruning step.
Pre-pruning does not affect the output of the pruning step ($\tilde A_t$);
the purpose is solely to make the pruning step faster.
One way to obtain the auxiliary abstractions is via
compositions with another abstraction $\hclass$: $\beta_t = \alpha_t \circ \hclass$.
Intuitively, $\hclass$ should neither be too coarse nor too fine.
If $\hclass$ is no abstraction, then pre-pruning is equivalent to running the pruning step;
if $\hclass$ is the trivial abstraction, pre-pruning will be fast but nothing will be pre-pruned.
A rule of thumb is that $\hclass$ should be complementary to $\alpha_t$
(we will see examples in \refsec{abstractions}).

\refthm{algorithm} states that the \PR\ algorithm is both
sound and complete.  In other words, pruning has no impact on our answer to a
query---we just compute it more efficiently.
The proof is given in \refapp{proofs} and uses \refthm{master}.

\begin{theorem}[Correctness of the \PR\ algorithm]
\label{thm:algorithm}
At iteration $t$,
the incrementally pruned abstraction $A_t$ is equivalent
to the full abstraction $\alpha_t(X)$ (formally, $\bP(\alpha_t(X)) = \bP(A_t)$).
If the algorithm returns ``proven,'' then $\bP(X) = \emptyset$ (the query is actually false).
\end{theorem}

>>>>

%In this section, we show how we can use pruning to obtain a sound analysis which
%is complete with respect to a given abstraction.
We first describe the core
idea behind pruning (\refsec{oneStep}) and then show how it can be used in the
Prune-Refine algorithm (\refsec{algorithm}).

@@

The left-hand side of \refeqn{master} corresponds to running the analysis with respect to $\alpha$.
The right-hand side corresponds to first pruning the input tuples $X$
with $\beta$ and then running the analysis with $\alpha$.
The theorem states that the two procedures obtain identical results
(the right-hand side is sound and complete with respect to the left-hand side).
The significance of this is that the right-hand side is often much cheaper
to compute than the left-hand side.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{algorithm}{The Prune-Refine algorithm}

\begin{figure}[t]
\begin{center} \scalebox{0.95}{\framebox{ \begin{minipage}{3.3in} % Begin model
\begin{center} Prune-Refine algorithm \end{center}
{\bf Input}: \\
\ind Sequence of abstractions: $\alpha_0 \preceq \alpha_1 \preceq \alpha_2 \preceq \cdots$ \\
\ind [Auxiliary abstractions: $\beta_t \preceq \alpha_t, t = 0, 1, 2, \dots$] \\
\ind $A_0 = \alpha_0(X)$, set of tuples \\
\\
For $t = 0, 1, 2, \dots$: \\
\ind [{\bf Pre-prune}: $A_t' \leftarrow A_t \cap \alpha_t(\bP(\beta_t(A_t)))$] \\
\ind {\bf Prune}: $\tilde A_t = \bP(A_t')$. If $\tilde A_t = \emptyset$: return {\em proven}. \\
\ind {\bf Refine}: $A_{t+1} = \alpha_{t+1}(\tilde A_t)$. \\
%\ind\ind\ind\ind\ind If $A_{t+1} = A_t$: return {\em not proven}.
\end{minipage} }} \end{center} % End model
\includegraphics[scale=0.5]{figures/algorithm}
\caption{\label{fig:pseudocode} The pseudocode and the schema for the Prune-Refine algorithm.
The algorithm maintains a set of (abstract) input tuples which could be
involved in some derivation of the query $\xo$ and attempts to prune down this set.
The basic version of the algorithm, which excludes the lines in square brackets and the dotted boxes,
simply alternates between pruning and refining.  The full version includes
a pre-pruning step, which uses auxiliary abstractions to further reduce the
number of tuples.
}
\end{figure}

We now turn \refthm{master} into a full algorithm, which we call the
Prune-Refine algorithm.  \reffig{pseudocode} shows the pseudocode
of the algorithm and a diagram showing the computation of the various abstract
input tuples computed.

% Basic
This algorithm essentially applies \refeqn{master} repeatedly.
We first present a simplified version of the algorithm which
ignores the pre-pruning step (we take $A_t = A_t'$).
We are given a sequence of successively finer abstractions $\alpha_0, \alpha_1, \dots$
(e.g., $\alpha_t = \pi_t$ for the $k$-limited abstractions)
and an input set of abstract tuples $A_0$, which is computed
under the initial abstraction $\alpha_0$.  Then the algorithm alternates between a pruning step
and a refining step, always maintaining only the abstract tuples
that could participate in a derivation of the query tuple $\xo$.
On iteration $t$, our current input tuples $A_t$ are first pruned to $\tilde A_t$ using $\bP$;
this is subsequently refined to $A_{t+1}$.
% Example
\reffig{graphIters} shows an example of running this algorithm on the graph
example from \reffig{graphExample}; the first pruning step is shown in
\reffig{graphDerivation}.

\Fig{figures/graphIters}{0.3}{graphIters}{
The abstract input tuples computed by
the Prune-Refine algorithm on the graph example from \reffig{graphExample} (without pre-pruning).
We are using $k$-limited abstractions ($\alpha_t = \pi_t$).
During the first pruning step, $\ext(2,\vec{0}*,\vec{2}*)$ is pruned from $A_0$,
yielding $\tilde A_0$.  In the refinement step, we expand $\vec{1}*$ to $\{ \vec{1} \}$ and $\vec{10}*$.
In the second pruning step, we prove the query
(pruning everything).
}

% Pre-pruning
Now we discuss pre-pruning.  Pre-pruning requires the user to provide
another sequence of successively finer abstractions $\beta_0, \beta_1, \dots$
which are coarser than $\alpha_0, \alpha_1, \dots$, respectively.
These abstractions will also be used to prune the input tuples.
The idea is that before refining $A_t$ to $A_{t+1}$,
we perform two steps of pruning:
(i) first we use $\beta_t$ in a pre-pruning step;
(ii) then we use $\alpha_t$ during the main pruning step.
Pre-pruning does not affect the output of the pruning step ($\tilde A_t$);
the purpose is solely to make the pruning step faster.

Pre-pruning requires a sequence of auxiliary abstractions $(\beta_t)$ which are coarser than the main abstractions $(\alpha_t)$.
We can obtain auxiliary abstractions by composing the main abstractions
with another abstraction $\hclass$ ($\beta_t = \alpha_t \circ \hclass$).
We can use any $\hclass$ for which $\alpha_t \circ \hclass$ yields a valid abstraction,
but the speedup we obtain from pre-pruning depends on the relationship between $\hclass$ and $\alpha_t$.
If $\hclass$ is the total abstraction ($\hclass(c) = \sC$), then pre-pruning will be fast but nothing will be pre-pruned,
so we get no speedup.
If $\hclass$ is no abstraction ($\hclass(c) = c$), then pre-pruning is equivalent to just running the pruning step,
so we again get no speedup.
A good rule of thumb is that $\hclass$ should be ``complementary'' to $\alpha_t$
(we will see some examples in \refsec{abstractions}).

\refthm{algorithm} states that the Prune-Refine algorithm is both
sound and complete.  In other words, pruning has no impact on our answer to a
query---we just compute it more efficiently.
The proof is given in \refapp{proofs} and uses \refthm{master}.

\begin{theorem}[Correctness of the Prune-Refine algorithm]
\label{thm:algorithm}
At iteration $t$,
the incrementally pruned abstraction $A_t$ is equivalent
to the full abstraction $\alpha_t(X)$ (formally, $\bP(\alpha_t(X)) = \bP(A_t)$).
If the algorithm returns ``proven,'' then $\bP(X) = \emptyset$ (the query is actually false).
\end{theorem}
